---
title: "Danish Kings"
author: "Adela Sobotkova"
date: "2025-01-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


The task here is to load your Danish Monarchs csv into R using the tidyverse toolkit, calculate and explore the kings' duration of reign and plot it over time.

## Load the kings

Make sure to first create an .Rproj workspace with a data/ folder where you place either your own dataset or the provided ‘kings’ dataset. 

1. Look at the dataset that are you loading and check what are its columns separated by? (hint: open it in plain text editor to see)

Your answer:

2. Create a `kings` object with different functions and inspect the different outputs.
Which of these functions is a `tidyverse` function? What is the `class()` of the ‘kings’ object and how many columns does the object have when created with these different functions?

-  `read.csv()`
-  `read_csv()`
-  `read.csv2()` 
-  `read_csv2()`
```{r eval=FALSE}
# FILL IN THE CODE BELOW and review the outputs
kings <- read.csv("data/_______")

kings <- read_csv("data/_______")

kings <- read.csv2("data/_______")

kings <- read_csv2( "data/_______")
```

Which function produced a nice tidy `tibble` output for your dataset? Show the dataset here:

```{r}
# YOUR CODE
```


## Calculate the duration of reign for all the kings in your table
You can calculate the duration of reign in years by subtracting the equivalents of your `startReign` from `endReign` columns and writing the result to a new column called `duration`. But first you need to check a few things:

* Is your data messy? Fix it before re-importing to R
* Do your start and end of reign columns contain NAs? Choose the right strategy to deal with them: `na.omit()`, `na.rm=TRUE`, `!is.na()`


Calculate the `duration` column in the kings dataset. You can use either the `$` in kings to access/create columns, or utilize the `mutate()` function from tidyverse. Check with your group to brainstorm the options.

```{r}
# YOUR CODE
```

## Calculate the average duration of reign for all rulers
Do you remember how to calculate an average on a vector object? If not, review the last two lessons and remember that a column is basically a vector. Next, practice creating tidyverse summaries or subsetting; both allow you to work with specific columns and do calculations on them.

* You first need to know how to select the relevant `duration` column before you run a `mean()` on it.
* Is your `duration` column a tibble or a vector (=part of dataframe)? The `mean()` function can only be run on a vector. The `summarize()` function works on a tibble.
* Are you getting an error that there are characters in your column? Coerce your data to numbers with `as.numeric()`.
* Remember to handle NAs: `mean(X, na.rm=TRUE)`


```{r}
# YOUR CODE
```

## How many days did the three longest-ruling monarchs rule? 

* Sort kings by reign `duration` in the descending order. Select the three longest-ruling monarchs with the `slice()` function
* Use `mutate()` to create `Days` column where you calculate the total number of days they ruled
* BONUS: consider the transition year (with 366 days) in your calculation!

```{r}
# YOUR CODE
```


## Plot the kings' duration of reign through time

What is the long-term trend in the duration of reign among Danish monarchs? How does it relate to the historical violence trends ? 

* Try to plot the duration of reign column in `ggplot` with `geom_point()` and  `geom_smooth()`
* In order to peg the duration (which is between 1-99) somewhere to the x axis with individual centuries, I recommend creating a new column `midyear` by subtracting the endYear from the startYear and dividing the result by two.
* Now you can plot the kings dataset, plotting `midyear` along the x axis and 'Duration' along y axis
* BONUS: add a title, nice axis labels to the plot and make the theme B&W and font bigger to make it nice and legible!

```{r}
# YOUR CODE
```

And to submit this rmarkdown, knit it. But first, clean up the code chunks, adjust the date, rename the author and change the `eval=FALSE` flag to `eval=TRUE` so your script generates a nice output.
Well done!

Well done!
